---
title: "Practical session on CO2 dataset"
author: "Florent Chuffart - florent.chuffart@univ-grenoble-alpes.fr"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---



```{r, echo=FALSE, eval=TRUE, label="loading libraries"}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=FALSE, results="hide")
```
# Loading data

```{r echo=TRUE, results="verbatim"}
data(CO2) # charge le jeu de données CO2 interne à R
head(CO2)
attach(CO2) # attachement -> pour acceder directement aux variables internes
names(CO2) <- c("ID_Plante", "Type_Plante", "Traitement", "C_CO2", "Assimilation") # renomage des variables
CO2.q <- CO2[CO2$Type_Plante=="Quebec",] # données restreintes au quebec
CO2.m <- CO2[CO2$Type_Plante=="Mississippi",] # données restreintes au mississippi

col.v <- c("black", "gray50") # vecteur de couleurs
head(CO2)
```



# Graphical Analysis

```{r echo=TRUE, results="verbatim"}
source("MSS2_TD1/TIS4_TD_CodeGraph1.R")
```

# ANOVA / linear regression

Many approachs:

  - linear regression, but data are note linear (saturation of CO2 assimilation)
  - linear regression on linear part of the curve (but we do not used the entire data set)
  - linear regression on transformed data (ex: log(assimilitaion)), but finding good transformation is not easy
  - ANOVA, but it do not take into account the order of the [CO2]

## ANOVA
  
```{r echo=TRUE, results="verbatim"}
m1 = lm(Assimilation~Type_Plante+Traitement+C_CO2  , data=CO2)
anova(m1)
summary(m1)

CO2$C_CO2_fact = as.factor(CO2$C_CO2)
m2 = lm(Assimilation~Type_Plante+Traitement+C_CO2_fact  , data=CO2)
anova(m2)
summary(m2)
```

Even if m1 is more free than m2 (df1 > df2), 
m1 is less powerfull than m2 (Rsquare2 > Rsquare1).

# Logictic regression

```{r echo=TRUE, results="verbatim"}
```

# Session Information

```{r, results="verbatim"}
sessionInfo()
```

