---
title: "Practical session on CO2 dataset"
author: "Florent Chuffart & Nicolas Glade"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---



```{r, echo=FALSE, eval=TRUE, label="loading libraries"}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=FALSE, results="hide")
```
# Loading data

```{r echo=TRUE, results="verbatim"}
data(CO2) # charge le jeu de données CO2 interne à R
head(CO2)
attach(CO2) # attachement -> pour acceder directement aux variables internes
names(CO2) <- c("ID_Plante", "Type_Plante", "Traitement", "C_CO2", "Assimilation") # renomage des variables
CO2.q <- CO2[CO2$Type_Plante=="Quebec",] # données restreintes au quebec
CO2.m <- CO2[CO2$Type_Plante=="Mississippi",] # données restreintes au mississippi

col.v <- c("black", "gray50") # vecteur de couleurs
head(CO2)
```

# Graphical Analysis (by Nicolas Glade)

```{r echo=TRUE, results="verbatim"}
source("MSS2_TD1/TIS4_TD_CodeGraph1.R")
detach(CO2)
```

# Modelizations

How *food* availability explains *growth* of plants in "Quebec" population?

```{r echo=TRUE, results="verbatim"}
d = data.frame(
  growth = CO2.q$Assimilation ,
  food =   CO2.q$C_CO2           
)
```

Several approaches:

  - linear regression, easy but data are not linear,
  - focus on linear part of the curve, it works but a part of the information is lost,
  - linear regression on transformed data, it is better but it‘s difficult to find a good transformation function, candidate are log transformation, single exponential model... ),
  - ANOVA, a part of the information is lost and the ANOVA has less degree of freedom, but with a good experimental design and enough replicates this model could better explain the data.

## Linear regression

$$growth \sim food$$
  
```{r echo=TRUE, results="verbatim"}
m = lm(growth ~ food, d)
```

Explained variance is poor.

```{r echo=TRUE, results="verbatim"}
summary(m)
```
Linear regression is not a good model, *growth* curve reached a plateau (fig).
Linear regression seems better with log transformation of *food*.

```{r echo=TRUE, results="verbatim"}
d$l2food = log2(d$food)
```


```{r}
# README RStudio config, uncheck: # preferences > R Markdown > show output inline for... 
layout(matrix(1:2, 1, byrow=TRUE), respect=TRUE)
plot(d$food, d$growth, main="growth~food")
## Model
# Y~X
# E(Y) = b.X
# E(Y) = b_0 + b_1.X
# Y_i = b_0 + b_1.X_i + e_i
m = lm(growth ~ food, d)
# summary(m)
# m$coefficients
# anova(m)
abline(a=m$coefficients[[1]], b=m$coefficients[[2]], col=2) # /!\ y = b.x + a
# residuals
# m$residuals
arrows(d$food, d$growth, d$food, d$growth-m$residuals, col=adjustcolor(4, alpha.f=0.5), length=0.1)
legend("topleft",c("regression line", "residuals"), col=c(2,4), lty=1, cex=.8)

plot(d$l2food, d$growth, main="growth~log(food)")
m = lm(growth ~ l2food, d)
abline(a=m$coefficients[[1]], b=m$coefficients[[2]], col=2) # /!\ y = b.x + a
arrows(d$l2food, d$growth, d$l2food, d$growth-m$residuals, col=adjustcolor(4, alpha.f=0.5), length=0.1)
legend("topleft",c("regression line", "residuals"), col=c(2,4), lty=1, cex=.8)
```

## Linear regression on log transformed data

$$growth \sim log(food)$$


Variance is better explained be this model.

```{r echo=TRUE, results="verbatim"}
summary(m)
```

## ANOVA 

In ANOVA, we consider *food* as a factor. 

The experimental design reveals 6 replicates per condition.

```{r echo=TRUE, results="verbatim"}
table(d$food)
```

$$growth \sim food$$

```{r}
d$ffood = as.factor(d$food)
layout(matrix(1:2, 1, byrow=TRUE), respect=TRUE)
p = plot(d$ffood, d$growth, main="growth~food", las=2)
m = lm(growth ~ ffood, d)
ffood = jitter(as.numeric(d$ffood))
arrows(ffood, d$growth, ffood, d$growth-m$residuals, col=adjustcolor(4, alpha.f=0.5), length=0.1)
legend("topleft","residuals", col=4, lty=1, cex=.8)
```

In our case, variance is fare better explained by the ANOVA.

```{r echo=TRUE, results="verbatim"}
summary(m)
```





# Prediction

'''{r eavl=FALSE}
CO2_train = CO2
CO2_test = CO2
m = glm(Type~Treatment+conc+uptake, CO2_train, family =binomial(link = 'logit'))
pred = predict.glm(m, newdata = CO2_test, type = "response")
plot(pred~CO2_test$Type)
pred_type = ifelse(pred>0.5, "Mississippi", "Quebec")
paste(round(sum(pred_type == CO2_test$Type) / nrow(CO2) * 100, 2), "% de bonne prédiction")


CO2$conc_fact = as.factor(CO2$conc)
CO2_train = CO2
CO2_test = CO2
m = glm(Type~Treatment+conc_fact+uptake, CO2_train, family =binomial(link = 'logit'))
pred = predict.glm(m, newdata = CO2_test, type = "response")
plot(pred~CO2_test$Type)
pred_type = ifelse(pred>0.5, "Mississippi", "Quebec")
paste(round(sum(pred_type == CO2_test$Type) / nrow(CO2) * 100, 2), "% de bonne prédiction")


CO2$conc_fact = as.factor(CO2$conc)
idx_train = sample(1:nrow(CO2), 63)
idx_test = setdiff(1:nrow(CO2), idx_train)
CO2_train = CO2[idx_train,]
CO2_test = CO2[idx_test,]
m = glm(Type~Treatment+conc+uptake, CO2_train, family =binomial(link = 'logit'))
pred = predict.glm(m, newdata = CO2_test, type = "response")
plot(pred~CO2_test$Type)
pred_type = ifelse(pred>0.5, "Mississippi", "Quebec")
paste(round(sum(pred_type == CO2_test$Type) / nrow(CO2_test) * 100, 2), "% de bonne prédiction")
'''

# Session Information

```{r, results="verbatim"}
sessionInfo()
```

